---
title: "Pub Quiz Stats"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10
)


if (!require(tidyverse)) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require(janitor)) {
  install.packages("janitor")
  library(janitor)
}
if (!require(lubridate)) {
  install.packages("lubridate")
  library(lubridate)
}
if (!require(tidytext)) {
  install.packages("tidytext")
  library(tidytext)
}
if (!require(ggplot2)) {
  install.packages("ggplot2")
  library(ggplot2)
}
if (!require(scales)) {
  install.packages("scales")
  library(scales)
}
if (!require(gt)) {
  install.packages("gt")
  library(gt)
}
if (!require(corrr)) {
  install.packages("corrr")
  library(corrr)
}
if (!require(googlesheets4)) {
  install.packages("googlesheets4")
  library(googlesheets4)
}
```

```{r data_import, include=FALSE}
sheet_url <- "https://docs.google.com/spreadsheets/d/12yUhkqq5ajM_uq7vLytH9cBceCCDFC0MEJluVX_aBdA/edit?usp=sharing"

# Sheet doesn't need auth as it has link editing enabled
gs4_deauth()

# Read the sheet and clean up the variable names
quiz_data <-
  read_sheet(sheet_url) %>%
  clean_names()
```

Report ran `r today()`.

```{r data_clean, include=FALSE}
# Set some vars for use with formatting the positions
position_order <- c(
  "winning_team",
  "second_place",
  "third_place",
  "fourth_place"
)
position_label <- c("1st", "2nd", "3rd", "4th")

min_quizzes <- 8
n_to_plot <- 12

clean_data <- quiz_data %>%
  rename(quiz_date = hosted_quiz) %>%
  mutate(
    quiz_date = as.Date(quiz_date),
    quiz_number = row_number(),
    time_weighting = (quiz_number * 2) %>% scale(center = FALSE)
  ) %>%
  # Restructure to make it easier to work with
  pivot_longer(
    cols = c(winning_team, second_place, third_place, fourth_place),
    names_to = "position",
    values_to = "team_members"
  ) %>%
  # Drop missing data
  drop_na() %>%
  separate_rows(team_members, sep = "-") %>%
  group_by(quiz_date) %>%
  mutate(team_number = row_number()) %>%
  ungroup() %>%
  # Restructure and tidy up the text
  unnest_tokens(
    output = "person",
    input = "team_members"
  ) %>%
  anti_join(get_stopwords(), by = c("person" = "word")) %>%
  # Fixes for names
  mutate(
    person = case_when(
      person == "cami" ~ "camille",
      person == "jonathan" ~ "jonty",
      person == "spud" ~ "emma",
      TRUE ~ person
    ),
    # Re-capitalise
    person = str_to_sentence(person)
  ) %>%
  # Create a nicely ordered factor
  mutate(position = ordered(position, position_order, position_label)) %>%
  # Calculate scaled scores
  group_by(quiz_date, team_number) %>%
  mutate(team_size = n()) %>%
  group_by(quiz_date) %>%
  # Score scaled according to the number of teams
  mutate(scaled_score = fct_rev(position) %>%
    as.integer() %>%
    scale(center = FALSE)) %>%
  # Scaled score inversely weighted by the size of the team
  # Also a cumulative mean of the weighted score
  mutate(
    team_weighted_score = scaled_score * max(team_size) / team_size,
    time_weighted_score = team_weighted_score * time_weighting
  ) %>%
  # Count quizzes per person
  group_by(person) %>%
  mutate(
    score_cummean = cummean(team_weighted_score),
    time_weighted_score_cummean = cummean(time_weighted_score),
    n_quizzes = n()
  ) %>%
  left_join(
    summarise(., n_quizzes = max(n_quizzes)) %>%
      mutate(n_quizzes_rank = rank(-n_quizzes, ties.method = "min"))
  ) %>%
  left_join(
    filter(., n_quizzes >= min_quizzes) %>%
      summarise(
        overall_rank = last(score_cummean),
        time_weighted_rank = last(time_weighted_score_cummean)
      ) %>%
      mutate(
        overall_rank = rank(-overall_rank, ties.method = "min"),
        time_weighted_rank = rank(-time_weighted_rank, ties.method = "min")
      )
  ) %>%
  filter(!(person %in% c("Calum", "Rory", "Jonty", "Ben", "Ian"))) %>%
  ungroup() %>%
  mutate(person = case_when(
    person == "Sarah" ~ "Sarah & Calum",
    person == "Viv" ~ "Viv & Rory",
    person == "Julia" ~ "Julia & Jonty",
    person == "Cathy" ~ "Cathy & Ben",
    person == "Kirsty" ~ "Kirsty & Ian",
    TRUE ~ person
  ))
```

## Hosting Summary

```{r host_table}
quiz_data %>%
  select(quizmasters, hosted_quiz, winning_team) %>%
  drop_na() %>%
  group_by(quizmasters) %>%
  summarise(
    n_hosted = n(),
    last_hosted = time_length(interval(max(hosted_quiz), today()), "weeks")
  ) %>%
  arrange(-n_hosted, last_hosted) %>%
  mutate(last_hosted = if_else(last_hosted == 1,
    str_glue("{ceiling(last_hosted)} week ago"),
    str_glue("{ceiling(last_hosted)} weeks ago")
  )) %>%
  rename(
    "Quiz Hosts" = quizmasters,
    "Times hosted" = n_hosted,
    "Last hosted" = last_hosted
  ) %>%
  gt()
```

## Score summary

Rank is only calculated for people who have at least `r min_quizzes` recorded. Time weighted rank gives more importance to more recent quizzes (weighted linearly)

```{r summary_table}
clean_data %>%
  group_by(person) %>%
  summarise(across(
    c(
      time_weighted_rank,
      time_weighted_score_cummean,
      overall_rank,
      score_cummean,
      n_quizzes,
      time_weighted_score
    ),
    last
  )) %>%
  arrange(time_weighted_rank) %>%
  drop_na() %>%
  mutate(across(ends_with("_rank"), ordinal)) %>%
  rename(
    Name = person,
    "Time-weighted Rank" = time_weighted_rank,
    "Mean score (time-weighted)" = time_weighted_score_cummean,
    "Overall Rank" = overall_rank,
    "Mean score (overall)" = score_cummean,
    "Number of quizzes" = n_quizzes,
    "Last quiz score" = time_weighted_score
  ) %>%
  gt() %>%
  tab_footnote(
    footnote = "Score is the score per quiz (1st, 2nd etc.),
    scaled and then weighted by the team size",
    locations = cells_column_labels(
      columns = vars("Mean score (overall)", "Mean score (time-weighted)")
    )
  ) %>%
  fmt_number(columns = vars(
    "Mean score (overall)",
    "Mean score (time-weighted)",
    "Last quiz score"
  ))
```

## Plots

### Position plot

```{r bar_plot_n}
# Plot of People by number of times they have placed.
clean_data %>%
  ggplot(aes(x = reorder(person, time_weighted_rank), fill = position)) +
  geom_bar() +
  xlab("Person") +
  ylab("Number of quizes") +
  ggtitle(
    label = "Count of positions per person",
    subtitle = "Sorted by time-weighted rank"
  ) +
  scale_fill_brewer("Team Position",
    type = "seq",
    palette = "OrRd",
    direction = -1
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

```{r plot_formatting}

colour_scale <- scale_colour_brewer("Person",
  type = "qual",
  palette = "Set3"
)

colour_scale <- scale_color_manual("Person",
  values = ggsci::pal_simpsons()(n_to_plot)
)

library(randomcoloR)
n <- 13
palette <- scale_colour_manual(values = distinctColorPalette(n))

time_trend_x <-
  scale_x_date(
    breaks = breaks_width("week", offset = -1),
    labels = date_format("%d %b")
  )
```

### Weighted score over time

```{r time_score_over_time}
# Scaled score over time
clean_data %>%
  filter(n_quizzes_rank <= n_to_plot) %>%
  ggplot(aes(
    x = quiz_date, y = time_weighted_score,
    colour = reorder(person, time_weighted_rank)
  )) +
  geom_smooth(se = FALSE, size = 0.5) +
  xlab("Date of quiz") +
  ylab("Weighted score") +
  ggtitle(
    label = "Actual time-weighted score over time"
  ) +
  colour_scale +
  time_trend_x +
  theme_minimal() +
    theme(axis.text.x = element_text(angle = 45))
```

### Mean score over time

```{r mean_time_score_over_time}
# Scaled score over time
clean_data %>%
  filter(n_quizzes_rank <= n_to_plot) %>%
  ggplot(aes(
    x = quiz_date, y = time_weighted_score_cummean,
    colour = reorder(person, time_weighted_rank)
  )) +
  geom_smooth(se = FALSE, size = 0.5) +
  xlab("Date of quiz") +
  ylab("Weighted score") +
  ggtitle(
    label = "Cummulative mean time-weighted score over time"
  ) +
  colour_scale +
  time_trend_x +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

### Un-weighted score over time

```{r score_over_time}
# Scaled score over time
clean_data %>%
  filter(n_quizzes_rank <= n_to_plot) %>%
  ggplot(aes(
    x = quiz_date, y = team_weighted_score,
    colour = reorder(person, overall_rank)
  )) +
  geom_smooth(se = FALSE, size = 0.5) +
  xlab("Date of quiz") +
  ylab("Weighted score") +
  ggtitle(
    label = "Actual un-weighted score over time"
  ) +
  colour_scale +
  time_trend_x +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

## Team corrolation plot

Correlation plot showing how often people are on the same team. Dark orange is always on the same team (couples), dark purple is rarely / never on the same team. Ideally it would all be white(ish), except for couples.

This excludes people with fewer than `r min_quizzes` recorded.

```{r corr_plot}

clean_data %>%
  filter(n_quizzes >= min_quizzes) %>%
  arrange(person) %>%
  pivot_wider(
    id_cols = quiz_date,
    names_from = person,
    values_from = team_number
  ) %>%
  select(-quiz_date) %>%
  correlate() %>%
  rearrange() %>%
  shave() %>%
  rplot(colors = c("#998ec3", "white", "#f1a340")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
```
