---
title: "Pub Quiz Stats"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10
)


if (!require(tidyverse)) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require(janitor)) {
  install.packages("janitor")
  library(janitor)
}
if (!require(lubridate)) {
  install.packages("lubridate")
  library(lubridate)
}
if (!require(tidytext)) {
  install.packages("tidytext")
  library(tidytext)
}
if (!require(ggplot2)) {
  install.packages("ggplot2")
  library(ggplot2)
}
if (!require(scales)) {
  install.packages("scales")
  library(scales)
}
if (!require(gt)) {
  install.packages("gt")
  library(gt)
}
if (!require(corrr)) {
  install.packages("corrr")
  library(corrr)
}
if (!require(googlesheets4)) {
  install.packages("googlesheets4")
  library(googlesheets4)
}
```

```{r data_import, include=FALSE}
source("code/import_data.R")
```

Report ran `r stamp("Monday 1 January 2020")(today())`.

```{r clean_data, include=FALSE}
# Set some vars for use with formatting the positions
position_order <- c(
  "winning_team",
  "second_place",
  "third_place",
  "fourth_place"
)

position_label <- c("1st", "2nd", "3rd", "4th")

min_quizzes <- 8
n_to_plot <- 14

source("code/clean_data.R")

# Combine couples who have always been together
clean_data <- clean_data %>% 
  filter(!(person %in% c("Calum", "Rory", "Jonty", "Ben", "Ian"))) %>%
  ungroup() %>% 
  mutate(person = case_when(
    person == "Sarah" ~ "Sarah & Calum",
    person == "Viv" ~ "Viv & Rory",
    person == "Julia" ~ "Julia & Jonty",
    person == "Cathy" ~ "Cathy & Ben",
    person == "Kirsty" ~ "Kirsty & Ian",
    TRUE ~ person
  ))

last_winning_team <- clean_data %>%
  filter(quiz_number == max(quiz_number), position == "1st")

pretty_last_winning_team <- last_winning_team %>%
  unnest_tokens(input = "person", output = "person") %>%
  pull(person) %>%
  str_to_sentence() %>%
  str_sort() %>%
  str_flatten(", ") %>%
  str_replace(", (\\w+?)$", " and \\1")

source("code/compute_streaks.R")
```

## Hosting Summary

There have been `r max(clean_data$quiz_number)` quizzes recorded since  the first (`r stamp("1 Jan 2020")(min(clean_data$quiz_date))`). The most recent recorded quiz was on `r stamp("Monday 1 January 2020")(last(clean_data$quiz_date))`, it was hosted by `r last(clean_data$quizmasters)` and the winning team was: `r pretty_last_winning_team`.

```{r host_table}
source("code/create_host_table.R")

hosted_table
```



## Score summary
<!-- TODO: Update this to deal with multiple people. -->
`r str_glue("{people_on_streak} is on a streak of {max_streak} wins {emo::ji('party')}")`


Rank is only calculated for people who have at least `r min_quizzes` recorded. Time weighted rank gives more importance to more recent quizzes (weighted linearly)

```{r summary_table}
clean_data %>%
  group_by(person) %>%
  summarise(across(
    c(
      time_weighted_rank,
      time_weighted_score_cummean,
      overall_rank,
      score_cummean,
      n_quizzes,
      time_weighted_score
    ),
    last
  )) %>%
  arrange(time_weighted_rank) %>%
  drop_na() %>%
  mutate(across(ends_with("_rank"), ordinal)) %>%
  rename(
    Name = person,
    "Time-weighted Rank" = time_weighted_rank,
    "Mean score (time-weighted)" = time_weighted_score_cummean,
    "Overall Rank" = overall_rank,
    "Mean score (overall)" = score_cummean,
    "Number of quizzes" = n_quizzes,
    "Last quiz score" = time_weighted_score
  ) %>%
  gt() %>%
  tab_footnote(
    footnote = "Score is the score per quiz (1st, 2nd etc.),
    scaled and then weighted by the team size",
    locations = cells_column_labels(
      columns = vars("Mean score (overall)", "Mean score (time-weighted)")
    )
  ) %>%
  fmt_number(columns = vars(
    "Mean score (overall)",
    "Mean score (time-weighted)",
    "Last quiz score"
  ))
```

## Position plot

```{r bar_plot_n}
# Plot of People by number of times they have placed.
clean_data %>%
  ggplot(aes(x = reorder(person, time_weighted_rank), fill = position)) +
  geom_bar() +
  xlab("Person") +
  ylab("Number of quizes") +
  ggtitle(
    label = "Count of positions per person",
    subtitle = "Sorted by time-weighted rank"
  ) +
  scale_fill_brewer("Team Position",
    type = "seq",
    palette = "OrRd",
    direction = -1
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

## Team corrolation plot

Correlation plot showing how often people are on the same team. Dark orange is always on the same team (couples), dark purple is rarely / never on the same team. Ideally it would all be white(ish), except for couples.

This excludes people with fewer than `r min_quizzes` recorded.

```{r corr_plot}

clean_data %>%
  filter(n_quizzes >= min_quizzes) %>%
  arrange(person) %>%
  pivot_wider(
    id_cols = quiz_date,
    names_from = person,
    values_from = team_number
  ) %>%
  select(-quiz_date) %>%
  correlate() %>%
  rearrange() %>%
  shave() %>%
  rplot(colors = c("#998ec3", "white", "#f1a340")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
```

## Time trends

```{r plot_formatting}

colour_scale <- scale_colour_brewer("Person",
  type = "qual",
  palette = "Set3"
)

colour_scale <- scale_color_manual("Person",
  values = ggsci::pal_simpsons()(n_to_plot)
)

library(randomcoloR)
n <- 13
palette <- scale_colour_manual(values = distinctColorPalette(n))

shape_scale <- scale_shape_manual("Person", values = rep(c(15, 16, 17), n_to_plot)[1:n_to_plot])

time_trend_x <-
  scale_x_date(
    breaks = breaks_width("week", offset = -1),
    labels = date_format("%d %b")
  )
```

### Weighted score over time

```{r time_score_over_time}
# Scaled score over time
clean_data %>%
  filter(n_quizzes_rank <= n_to_plot) %>%
  ggplot(aes(
    x = quiz_date, y = time_weighted_score,
    colour = reorder(person, time_weighted_rank)
  )) +
  geom_smooth(se = FALSE, size = 0.5) +
  xlab("Date of quiz") +
  ylab("Weighted score") +
  ggtitle(
    label = "Actual time-weighted score over time"
  ) +
  colour_scale +
  time_trend_x +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

### Mean score over time

```{r mean_time_score_over_time}
# Scaled score over time
clean_data %>%
  filter(n_quizzes_rank <= n_to_plot) %>%
  ggplot(aes(
    x = quiz_date, y = time_weighted_score_cummean,
    colour = reorder(person, time_weighted_rank)
  )) +
  geom_smooth(se = FALSE, size = 0.5) +
  xlab("Date of quiz") +
  ylab("Weighted score") +
  ggtitle(
    label = "Cummulative mean time-weighted score over time"
  ) +
  colour_scale +
  time_trend_x +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

### Un-weighted score over time

```{r score_over_time}
# Scaled score over time
clean_data %>%
  filter(n_quizzes_rank <= n_to_plot) %>%
  ggplot(aes(
    x = quiz_date, y = team_weighted_score,
    colour = reorder(person, overall_rank),
    shape = reorder(person, overall_rank)
  )) +
  geom_line(size = 0.1, linetype = "dotted", lineend = "round") +
  geom_point() +
  xlab("Date of quiz") +
  ylab("Weighted score") +
  ggtitle(
    label = "Actual un-weighted score over time"
  ) +
  colour_scale +
  shape_scale +
  time_trend_x +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```
### Positions over time


```{r position_over_time}
clean_data %>%
  filter(n_quizzes_rank <= n_to_plot) %>%
  ggplot(aes(
    x = quiz_date, y = fct_rev(position),
    colour = reorder(person, overall_rank),
    shape = reorder(person, overall_rank)
  )) +
  geom_jitter(width = 0, height = 0.25) +
  xlab("Date of quiz") +
  ylab("Team position") +
  ggtitle(
    label = "Positions over time"
  ) +
  colour_scale +
  shape_scale +
  time_trend_x +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```
